---
title: "PracticalMachineLearning"
author: "ArjunVenkat"
date: "October 24, 2015"
output: html_document
---
##Practical Machine Learning 
#Prediction Assignment Writeup
This document describe the analysis done for the prediction assignment of the practical machine learning course.

##Required Libraries
The packages used in the analysis are as follows:
```{r}
options(warn=-1)
library(caret)
library(randomForest)
library(Hmisc)
library(foreach)
library(doParallel)
set.seed(4356)
```
The first step is to load the csv file data to dataframe and analyze the type & the completion rate of the data.
```{r}
data <- read.csv("pml-training.csv")
summary(data)
describe(data)
sapply(data, class)
str(data)
```
##Reading the data

Read the training data into R, identifying "NA", "" and "#DIV/0!" as NA strings
```{r}
data <- read.csv("pml-training.csv", na.strings=c("#DIV/0!") )
```
And force the cast to numeric values for the specified columns 
```{r}
cData <- data
for(i in c(8:ncol(cData)-1)) {cData[,i] = as.numeric(as.character(cData[,i]))}
```
To manage the second issue we will select as feature only the column with a 100% completion rate We will also eliminate some features like "X"", timestamps, "new_window" and "num_window". 

```{r}
featuresnames <- colnames(cData[colSums(is.na(cData)) == 0])[-(1:7)]
features <- cData[featuresnames]
```

Spliting the data into training and test sets
```{r}
xdata <- createDataPartition(y=features$classe, p=3/4, list=FALSE )
training <- features[xdata,]
testing <- features[-xdata,]
```
Now we gonna process 4 random forest with 150 trees each and combine then to have a random forest model with a total of 600 trees.
```{r}
registerDoParallel()
model <- foreach(ntree=rep(150, 4), .combine=randomForest::combine, .packages='randomForest') %dopar% randomForest(training[-ncol(training)], training$classe, ntree=ntree)
```
To evaluate the model we will use the confusionmatrix method and we will focus on accuracy, sensitivity & specificity metrics :
```{r}
predictionsTr <- predict(model, newdata=training)
confusionMatrix(predictionsTr,training$classe)


predictionsTe <- predict(model, newdata=testing)
confusionMatrix(predictionsTe,testing$classe)
```

##Conclusion
As seen by the result of the confusionmatrix, the model is good and efficient because it has an accuracy of 0.997 and very good sensitivity & specificity values on the testing dataset.Also i accumulated 20/20 in Coursera Assignment Submission which proves it efficiency.